A DPG method for steady viscous compressible flow

### 2013-07-22 docs/Jesse/NavierStokes paper/NonlinDPG.tex
updating confusion, dissertation, and NS

\section{DPG for nonlinear problems}

In this chapter, we extend DPG to the nonlinear setting and apply it to two problems in computational fluid dynamics.  We take as our starting point a nonlinear variational formulation
$b(u,v) = l(v)$, which is linear in $v$, but not in $u$.  An appropriate linearization gives
\[
b_u(\Delta u,v) = l(v) - b(u,v),
\]
where $b_u(\Delta u,v)$ is the linearization of $b(u,v)$ with respect to $u$.  Let $B(u)$ and $B_u\Delta u$ be the variational operators associated with $b(u,v)$ and $b_u(\Delta u,v)$, respectively.  We define two additional measures: 
\begin{align*}
\|\Delta u\|_E \coloneqq & \nor{B_u\Delta u}_{V'} = \nor{R_V^{-1} B_u\Delta u}_V = \sup_{v\in V} \frac{b_u(\Delta u,v)}{\nor{v}_V}\\
\nor{R(u)}_E \coloneqq & \nor{B(u)-l}_E = \nor{B(u) - l}_{V'}  = \nor{R_V^{-1} B(u)-l}_V = \sup_{v\in V} \frac{b(u,v)-l(v)}{\nor{v}_V}
\end{align*}
These two quantities are measures of the linearized update $\Delta u$ and the nonlinear residual in the appropriate norm in the dual space $V'$.  The first will be used to measure convergence of a nonlinear solution scheme to a stable discrete solution, while the second will be used to assess the convergence of the discrete solution to the continuous solution.

@@

The solution of a nonlinear problem is most commonly found using an iterative method, where a series of solutions to linear problems is expected to converge to the nonlinear solution.  We use two main methods to iterate to a nonlinear solution.  
\begin{itemize}
\item{} \textbf{(Damped) Newton iteration :} Given the linearized system $b_u(\Delta u,v) = b(u,v)-l(v)$, we begin with some initial guess $u\coloneqq u_0$ and solve for $\Delta u_0$.  The process is then repeated with $u\coloneqq u_{i+1} \coloneqq u_i + \alpha_i \Delta u_i$, where $\alpha_i \in (0,1]$ is some damping parameter that may limit the size of the Newton step in order to optimize the rate of convergence or preserve physicality of the solution.  The solution is considered converged when $\nor{\Delta u}_E \leq tol$.  
\item{} \textbf{Pseudo-time stepping: } An alternative method for the solution of steady-state systems is to use a pseudo-timestepping method.  The most common approach is to discretize the equations in time using a stable, implicit method --- if $Lu = f$ is our nonlinear problem and $L_u$ is the linearization of the nonlinear operator $L$ with respect to $u$, then the pseudo-timestepping method solves at each discrete time $t_{i}$
\[
\pd{u}{t} + Lu = f \rightarrow \frac{u(t_i) - u(t_{i-1})}{\Delta t} + L_{u(t_i)}\Delta u(t_i) = (f - Lu(t_i)).
\]
The solution at the next timestep is then set $u(t_{i+1}) \coloneqq u(t_i) + \Delta u(t_i)$.  This procedure is then repeated for the next timestep $t_{i+1}$ until the transient residual decreases such that $\nor{u(t_i) - u(t_{i-1})}_{L^2(\Omega)} = \nor{\Delta u(t_i)}_{L^2(\Omega)} \leq tol$.\footnote{Strictly speaking, seeking the solution at each timestep involves the solution of a nonlinear problem, requiring a Newton-type iteration to solve for $u(t_i)$.  However, for most applications, it is sufficient to approximate the nonlinear solution using a single Newton solve at each time step.} 
\end{itemize}

In practice, most compressible flow solvers opt for the pseudo-time step method over the direct Newton iteration due to the difficulty of convergence and sensitivity of the Newton iteration to initial guess\cite{libMeshPaper}.  Though the convergence of the pseudo-time step is slower, the addition of the zero-order transient terms ``regularizes" the problem and makes it less difficult to solve.\footnote{The addition of a zero-order term ``regularizes" an equation by adding to it a positive-definite $L^2$ projection operator.  In the limit as $\Delta t\rightarrow 0$, the solution at $t_i$ will simply return the $L^2$ projection of the solution at the previous timestep.}

A second class of nonlinear solvers are optimization methods.  Since DPG allows for the formulation of a discrete nonlinear residual, it is possible to formulate the nonlinear DPG problem as a minimization problem and use an optimization method to solve the discrete nonlinear problem.  This approach has been successfully implemented by Peraire et al.\ in solving compressible gas dynamics problems on uniform grids using a modified version of the ultra-weak variational formulation \cite{MITDPG}.  An additional advantage of such an approach would be the more direct enforcement of physical constraints, which are treated in an ad-hoc manner in most compressible Navier-Stokes solvers.

@@

A recent theoretical development is the formulation of a DPG method that aims to minimize a nonlinear residual. Given two Hilbert spaces --- a trial space $U$ and test space $V$ --- our nonlinear variational formulation can be written as $b(u,v) = l(v)$, with the corresponding operator form of the formulation in $V'$
\[
B(u) = l.
\]
We can apply the steps used to derive the DPG method for linear problems to the nonlinear setting as well. Given a finite dimensional subspace $U_h \subset U$, we consider the discrete nonlinear residual
\[
J(u_h) \coloneqq \frac{1}{2}\nor{R_V^{-1}\left(B(u_h)-l\right)}_V^2.
\]
Our goal is to solve
\[
u_h = \arg \min_{w_h\in U_h}J(w_h).
\]
Similarly to the linear case, we can take the Gate\'aux derivative to arrive at a necessary condition for $u_h$ to minimize $J(u_h)$ 
\begin{align*}
\LRa{J'(u_h), \delta u_h} &= \LRp{R_V^{-1}\left(B(u_h)-l\right),R_V^{-1}B'(u_h;\delta u_h)}_V, \quad \delta u_h \in U_h. 
\end{align*}
As the above is a nonlinear equation, we seek its solution through linearization. Differentiating a second time in $u$, we arrive at
\begin{align*}
\LRa{J''(u_h), \Delta u_h} =& \LRa{B'(u_h;\Delta u_h), B'(u_h;\delta u_h)}_V \\
& + \LRa{\left(B(u_h)-l\right),B''(u_h;\delta u_h,\Delta u_h)}_V \\
=& \LRp{R_V^{-1}B'(u_h;\Delta u_h), R_V^{-1}B'(u_h;\delta u_h)}_V \\
& + \LRp{R_V^{-1}\left(B(u_h)-l\right),R_V^{-1}B''(u_h;\delta u_h,\Delta u_h)}_V 
\end{align*}
where $B''(u_h;\delta u_h,\Delta u_h)$ denotes the Hessian of $B(u_h)$, evaluated using both $\delta u_h$ and $\Delta u_h$.

@@

>>>>

\section{DPG for nonlinear problems}
\seclab{sec:nonlinDPG}
In this chapter, we extend DPG to the nonlinear setting and apply it to two problems in computational fluid dynamics.  We take as our starting point a nonlinear variational formulation
$b(u,v) = l(v)$, which is linear in $v$, but not in $u$.  An appropriate linearization gives
\[
b_u(\Delta u,v) = l(v) - b(u,v),
\]
where $b_u(\Delta u,v)$ is the linearization of $b(u,v)$ with respect to $u$.  Let $B(u)$ and $B_u\Delta u$ be the variational operators associated with $b(u,v)$ and $b_u(\Delta u,v)$, respectively.  We define two additional measures: 
\begin{align*}
\|\Delta u\|_E \coloneqq & \nor{B_u\Delta u}_{V'} = \nor{R_V^{-1} B_u\Delta u}_V = \sup_{v\in V} \frac{b_u(\Delta u,v)}{\nor{v}_V}\\
\nor{R(u)}_E \coloneqq & \nor{B(u)-l}_E = \nor{B(u) - l}_{V'}  = \nor{R_V^{-1} B(u)-l}_V = \sup_{v\in V} \frac{b(u,v)-l(v)}{\nor{v}_V}
\end{align*}
These two quantities are measures of the linearized update $\Delta u$ and the nonlinear residual in the appropriate norm in the dual space $V'$.  The first will be used to measure convergence of a nonlinear solution scheme to a stable discrete solution, while the second will be used to assess the convergence of the discrete solution to the continuous solution.

@@

The solution of a nonlinear problem is most commonly found using an iterative method, where a series of solutions to linear problems is expected to converge to the nonlinear solution.  We use two main methods to iterate to a nonlinear solution.  
\begin{itemize}
\item{} \textbf{(Damped) Newton iteration :} Given the linearized system $b_u(\Delta u,v) = b(u,v)-l(v)$, we begin with some initial guess $u\coloneqq u_0$ and solve for $\Delta u_0$.  The process is then repeated with $u\coloneqq u_{i+1} \coloneqq u_i + \alpha_i \Delta u_i$, where $\alpha_i \in (0,1]$ is some damping parameter that may limit the size of the Newton step in order to optimize the rate of convergence or preserve physicality of the solution.  The solution is considered converged when $\nor{\Delta u}_E \leq tol$.  
\item{} \textbf{Pseudo-time stepping: } An alternative method for the solution of steady-state systems is to use a pseudo-timestepping method.  The most common approach is to discretize the equations in time using a stable, implicit method --- if $Lu = f$ is our nonlinear problem and $L_u$ is the linearization of the nonlinear operator $L$ with respect to $u$, then the pseudo-timestepping method solves at each discrete time $t_{i}$
\[
\pd{u}{t} + Lu = f \rightarrow \frac{u(t_i) - u(t_{i-1})}{\Delta t} + L_{u(t_i)}\Delta u(t_i) = (f - Lu(t_i)).
\]
The solution at the next timestep is then set $u(t_{i+1}) \coloneqq u(t_i) + \Delta u(t_i)$.  This procedure is then repeated for the next timestep $t_{i+1}$ until the transient residual decreases such that $\nor{u(t_i) - u(t_{i-1})}_{L^2(\Omega)} = \nor{\Delta u(t_i)}_{L^2(\Omega)} \leq tol$.\footnote{Strictly speaking, seeking the solution at each timestep involves the solution of a nonlinear problem, requiring a Newton-type iteration to solve for $u(t_i)$.  For most applications, it is sufficient to approximate the nonlinear solution using a single Newton solve at each time step.} 
\end{itemize}

In practice, most compressible flow solvers opt for the pseudo-time step method over the direct Newton iteration due to the difficulty of convergence and sensitivity of the Newton iteration to initial guess\cite{libMeshPaper, gunzburger1989finite}.  Though the convergence of the pseudo-time step is slower, the addition of the zero-order transient terms ``regularizes" the problem and makes it less difficult to solve.\footnote{The addition of a zero-order term ``regularizes" an equation by adding to it a positive-definite $L^2$ projection operator.  In the limit as $\Delta t\rightarrow 0$, the solution at $t_i$ will simply return the $L^2$ projection of the solution at the previous timestep.}

A second class of nonlinear solvers are optimization methods.  Since DPG allows for the formulation of a discrete nonlinear residual, it is possible to formulate the nonlinear DPG problem as a minimization problem and use an optimization method to solve the discrete nonlinear problem.  This approach has been successfully implemented in solving compressible gas dynamics problems on uniform grids using a modified version of the ultra-weak variational formulation \cite{MITDPG, optDPG}.  An additional advantage of such an approach would be the more direct enforcement of physical constraints, which are treated in an ad-hoc manner in most compressible Navier-Stokes solvers.

@@

A recent theoretical development is the formulation of a DPG method that aims to minimize a nonlinear residual. Given two Hilbert spaces --- a trial space $U$ and test space $V$ --- our nonlinear variational formulation can be written as $b(u,v) = l(v)$, with the corresponding operator form of the formulation in $V'$
\[
B(u) = l.
\]
We can apply the steps used to derive the DPG method for linear problems to the nonlinear setting as well. Given a finite dimensional subspace $U_h \subset U$, we consider the discrete nonlinear residual
\[
J(u_h) \coloneqq \frac{1}{2}\nor{R_V^{-1}\left(B(u_h)-l\right)}_V^2.
\]
Our goal is to solve
\[
u_h = \arg \min_{w_h\in U_h}J(w_h).
\]
We note that, under convergence of any of the nonlinear iterations described above, the converged DPG solution does in fact satisfy the above minimization problem \cite{MITDPG}, and that the DPG energy error converges to the nonlinear residual.  However, it is theoretically possible to accelerate convergence of a nonlinear iteration to the minimizer $u_h$, using the same minimum residual framework DPG is based upon.

Similarly to the linear case, we can take the Gate\'aux derivative to arrive at a necessary condition for $u_h$ to minimize $J(u_h)$ 
\begin{align*}
\LRa{J'(u_h), \delta u_h} &= \LRp{R_V^{-1}\left(B(u_h)-l\right),R_V^{-1}B'(u_h;\delta u_h)}_V, \quad \delta u_h \in U_h. 
\end{align*}
As the above is a nonlinear equation, we seek its solution through linearization. Differentiating a second time in $u$, we arrive at
\begin{align*}
\LRa{J''(u_h), \Delta u_h} =& \LRa{B'(u_h;\Delta u_h), B'(u_h;\delta u_h)}_V \\
& + \LRa{\left(B(u_h)-l\right),B''(u_h;\delta u_h,\Delta u_h)}_V \\
=& \LRp{R_V^{-1}B'(u_h;\Delta u_h), R_V^{-1}B'(u_h;\delta u_h)}_V \\
& + \LRp{R_V^{-1}\left(B(u_h)-l\right),R_V^{-1}B''(u_h;\delta u_h,\Delta u_h)}_V 
\end{align*}
where $B''(u_h;\delta u_h,\Delta u_h)$ denotes the Hessian of $B(u_h)$, evaluated using both $\delta u_h$ and $\Delta u_h$.

@@

\subsection{DPG as a Gauss-Newton approximation}

The above Hessian-based nonlinear DPG method essentially replaces a simple linearization of with a higher-order expansion of the residual $J(u_h)$.  However, the application of DPG to the linearized equations actually yields a Gauss-Newton method for minimizing $J(u_h)$.

Recall that the discrete DPG method applied to a linear problem solves
\[
{B}^T{R_V}^{-1}{B}u - {B}^T{R_V}^{-1}l  = 0
\]
where ${B}$ is the rectangular matrix of size $m\times n$ resulting from the variational form $b(u,v)$ (where $m = \dim(U_h)$ and $n = \dim(V_h)$), and ${R_V}$ is the Gram matrix of dimension $n\times n$ resulting from choice of norm $\nor{\cdot}_V$ and inner product on $V$.  If we have the nonlinear equation ${B}(u_h)-{l}$ and linearize it, we have ${B}_u\triangle u_h = {r}(u_h)$, where $r(u_h) = l-B(u_h)$ is the residual.  The solution of the linearized equation by DPG minimizes then $\nor{r(u_h) - B_u\triangle u_h}_{V'}^2$, which we can recognize as linear least squares equations from which the Gauss-Newton method is derived \cite{Nocedal2006NO}.

Initial experiments with Hessian-based nonlinear DPG under Burgers' equation indicate that eigenvalues of the stiffness matrix can become negative if the Hessian is applied.  While the Gauss-Newton iteration may not converge as rapidly as the Hessian-based version of DPG, it does guarantee positive eigenvalues of the stiffness matrix, such that $\Delta u_h$ is always a descent direction.

### 2013-07-22 docs/Jesse/NavierStokes paper/outline.tex
updating confusion, dissertation, and NS

\documentclass{report}

\usepackage{utproposal}  		% Dissertation package style file.

\usepackage{array}
\usepackage{amsmath,amssymb,amsfonts,mathrsfs,amsthm}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{algorithm, algorithmic}
\usepackage{bm}
\usepackage{tikz}
\usepackage[normalem]{ulem}

\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{color}
\usepackage{undertilde}
\usepackage[colorlinks = true, filecolor = red, urlcolor = blue, linkcolor = black]{hyperref}
\usepackage{pdflscape}
\usepackage{pifont}
%\usepackage{fullpage}
\setlength\textwidth{6in}
\setlength\textheight{8in}
\setlength\oddsidemargin{0.25in} % LaTeX adds a default 1in to this!
\setlength\evensidemargin{0.25in}
\setlength\topmargin{-0.0in} % LaTeX adds a default 1in to this!
\setlength\headsep{0in}
\setlength\headheight{0in}
\setlength\footskip{1in}

@@

\newcommand{\circone}{\ding{192}}
\newcommand{\circtwo}{\ding{193}}
\newcommand{\circthree}{\ding{194}}
\newcommand{\circfour}{\ding{195}}
\newcommand{\circfive}{\ding{196}}

@@

@@

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\tanbui}[2]{\textcolor{blue}{\sout{#1}} \textcolor{red}{#2}}
\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

@@

%%%%%%%%%%%%%%% End def of new commands %%%%%%%%%%%%%%%%%%%

\author{Jesse Chan}
\address{910 E. 32nd St, Apt.\ 101\\ Austin, Texas 78705}  % Required

\title{A DPG method for compressible flow problems}

\supervisor
	[Leszek Demkowicz]
	{Robert Moser}

\committeemembers
	[Todd Arbogast]
	[Omar Ghattas]
	{Venkat Raman}

\previousdegrees{B.A.}

\oneandonehalfspacequote

\topmargin 0.125in

\makeindex

%%%%%%%%%%%%%%% Start of thesis %%%%%%%%%%%%%%%%%%%

\begin{document}
%\copyrightpage          % Produces the copyright page.
%\commcertpage           % Produces the Committee Certification
\titlepage

\tableofcontents   % Table of Contents will be automatically
                   % generated and placed here.

%\listoftables      % List of Tables and List of Figures will be placed
%\listoffigures     % here, if applicable.

\chapter{Introduction}

\input{proposalIntro}

\chapter{Range of problems}

\input{problemRange}

\chapter{Discontinuous Petrov-Galerkin: a minimum residual method for linear problems}

\input{DPGtheory}

\chapter{A robust DPG method for convection-diffusion}

\input{DPGrobustness}

\chapter{Proposed work}

\input{NonlinDPG}
\input{BurgersNotes}
\input{NSNotes}
\input{areaReq}

\bibliographystyle{plain}
\bibliography{CFD_intro,DPG_old,paper,LitRev,NSNotes}

%\chapter{Appendix}
\appendix
\input{paper_appendix.tex}
%\input{hyperbolicDPG.tex}

>>>>

%&pdflatex
\documentclass{elsarticle}

\usepackage{array}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{mathtools}
\usepackage{algorithm, algorithmic}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{color}
\usepackage{undertilde}
\usepackage[colorlinks = true, filecolor = red, urlcolor = blue, linkcolor = black]{hyperref}

\setlength\textwidth{6in}
\setlength\textheight{8in}
\setlength\oddsidemargin{0.25in} % LaTeX adds a default 1in to this!
\setlength\evensidemargin{0.25in}
\setlength\topmargin{-0.0in} % LaTeX adds a default 1in to this!
\setlength\headsep{0in}
\setlength\headheight{0in}
\setlength\footskip{1in}

@@

@@

\newtheorem{remark}{Remark}

@@

\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRb}[1]{\left| #1 \right|}

\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\LRs{\![#1]\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

@@

%%%%%%%%%%%%%%% End def of new commands %%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%% Start of thesis %%%%%%%%%%%%%%%%%%%

\begin{document}
\begin{frontmatter}

\author{Jesse Chan\corref{cor1}}
\ead{jchan@ices.utexas.edu}
\cortext[cor1]{Principal Corresponding Author, Institute for Computational Engineering and Sciences, The University of Texas at Austin}
\cortext[affil]{Institute for Computational Engineering and Sciences, The University of Texas at Austin}
\author{Leszek Demkowicz\corref{affil}}
\ead{leszek@ices.utexas.edu}
\author{Robert Moser\corref{affil}}
\ead{rmoser@ices.utexas.edu}

\title{A DPG method for steady viscous compressible flow}

\begin{abstract}
The Discontinuous Petrov-Galerkin (DPG) method is a class of novel higher order adaptive finite element methods derived from the minimization of the residual of the variational problem \cite{DPG2}, and has been shown to deliver a method for convection-diffusion that is provably robust in the diffusion parameter \cite{DPGrobustness,DPGrobustness2}.  In this work, the DPG method is extrapolated to nonlinear systems, and applied to several problems in fluid dynamics whose solutions exhibit boundary layers or singularities in stresses.  In particular, the effectiveness of DPG as a numerical method for compressible flow is assessed with the application of DPG to two model problems over a range of Mach numbers and laminar Reynolds numbers using automatic adaptivity for higher order elements, beginning with highly under-resolved coarse initial meshes.  
\end{abstract}

\end{frontmatter}

%\maketitle

%\tableofcontents

\input{paperIntro}

\input{DPGtheory}
\input{confusionSummary}

\input{NonlinDPG}
\input{BurgersNotes}
\input{NSNotes}

\input{Conclusions}

%\bibliographystyle{plain}
\bibliographystyle{unsrt}
\bibliography{paper,NSNotes,DPGarticles}

### 2013-07-22 docs/Jesse/NavierStokes paper/problemRange.tex
updating confusion, dissertation, and NS

\begin{figure}[!h]
\centering
\includegraphics[scale=.45]{figs/CFD_tree.pdf}
\caption{A diagram of common CFD problems and their simplifying assumptions.}
\end{figure}

@@

In 2D, the classical form of the Navier-Stokes equations involve the fluid density $\rho$, velocity in the $x$ and $y$ directions $u$ and $v$ (or $u_1$ and $u_2$), respectively, temperature $T$, energy per unit mass $e$, and stress and heat flux vectors $\boldsymbol \sigma_i$ and $\vec{q}$. The equations are as follows:
\begin{itemize}
\item{\textbf{Mass conservation}}
\begin{align*}
\pd{\rho}{t} + \div \vecttwo{\rho u }{\rho v} &= 0
\end{align*}

\item{\textbf{Momentum conservation}}
\begin{align*}
\pd{\rho u_1}{t} + \div \left(\vecttwo{\rho u^2+p }{\rho u v} - \boldsymbol \sigma_{1}\right) &=0\\
\pd{\rho u_2}{t} + \div \left(\vecttwo{\rho u v}{\rho v^2+p } - \boldsymbol \sigma_{2}\right) &=0
\end{align*}

\item{\textbf{Energy conservation}}
\begin{align*}
\pd{\rho e}{t} + \div \left(\vecttwo{((\rho e)+p)u}{((\rho e)+p)v} - \boldsymbol \sigma_1 \cdot \boldsymbol u- \boldsymbol \sigma_2 \cdot \boldsymbol u + \vec{q}\right) &=0
\end{align*}
\end{itemize}

@@

The scope of this dissertation will not deal with these two equations --- the Stokes equations are treated in \cite{stokesDPG}, and the incompressible Navier-Stokes are covered in an upcoming dissertation.

@@

The Burgers' equation is physically derived from the incompressible Navier-Stokes equations under the assumption that $\grad p \approx 0$, or that the pressure field is near constant. A feature of the Burgers' equation not present in convection-diffusion is that, due to the presence of the nonlinear term, it can develop shock discontinuities in its solutions in finite time.  The Burgers' equation has also been used to study the phenomenon of turbulence; however, the Burgers equation does not exhibit the chaotic nature and sensitivity to initial conditions that characterizes turbulence as observed in the full and incompressible Navier-Stokes equations.

The Burgers' equation is also the simplest nonlinear extension of the linear convection-diffusion equation, and exact solutions can sometimes be found using the method of characteristics. In the scope of this dissertation proposal, Burgers shall be used as to test the extension of our numerical method to nonlinear problems.

@@

\begin{figure}[!h]
\centering
\includegraphics[scale = .22]{figs/convCirc.png}
\caption{Setup for the vortex problem.}
\label{fig:convCirc}
\end{figure}
Formally speaking, the solution to the vortex problem is taken to be the solution to the convection-diffusion equation (with appropriate outflow boundary conditions) as the viscosity tends towards zero, in which case, the solution in the interior would be uniformly zero (this technique is referred to in mathematical literature as the ``vanishing viscosity" method, and is used to define unique solutions in the inviscid limit).  This motivates the need for \emph{artificial viscosity} methods with which to regularize inviscid solutions.  The topic is expansive, and we direct the reader towards \cite{Barter} for a more detailed discussion of past and present artificial viscosity methods.

>>>>

\begin{figure}[!h]
\centering
\includegraphics[scale=.475]{figs/CFD_tree.pdf}
\caption{A diagram of common CFD problems and their simplifying assumptions.}
\end{figure}

@@

In 2D, the classical form of the Navier-Stokes equations involve the fluid density $\rho$, velocity in the $x$ and $y$ directions $u_1$ and $u_2$, respectively, temperature $T$, energy per unit mass $e$, and stress and heat flux vectors $\boldsymbol \sigma_i$ and $\vec{q}$. The equations are as follows:
\begin{itemize}
\item{\textbf{Mass conservation}}
\begin{align*}
\pd{\rho}{t} + \div \vecttwo{\rho u_1 }{\rho u_2} &= 0
\end{align*}

\item{\textbf{Momentum conservation}}
\begin{align*}
\pd{\rho u_1}{t} + \div \left(\vecttwo{\rho u_1^2+p }{\rho u_1 u_2} - \boldsymbol \sigma_{1}\right) &=0\\
\pd{\rho u_2}{t} + \div \left(\vecttwo{\rho u_1 u_2}{\rho u_2^2+p } - \boldsymbol \sigma_{2}\right) &=0
\end{align*}

\item{\textbf{Energy conservation}}
\begin{align*}
\pd{\rho e}{t} + \div \left(\vecttwo{((\rho e)+p)u_1}{((\rho e)+p)u_2} - \boldsymbol \sigma_1 \cdot \boldsymbol u- \boldsymbol \sigma_2 \cdot \boldsymbol u + \vec{q}\right) &=0
\end{align*}
\end{itemize}

@@

The scope of this dissertation will not deal with these two equations --- the Stokes equations are treated in \cite{stokesDPG}, and the incompressible Navier-Stokes are covered in the upcoming dissertation of Nathan Roberts.

@@

The Burgers' equation is physically derived from the incompressible Navier-Stokes equations under the assumption that $\grad p \approx 0$, or that the pressure field is near constant. A feature of the Burgers' equation not present in convection-diffusion is that, due to the presence of the nonlinear term, it can develop shock discontinuities in its solutions in finite time.  The Burgers' equation has also been used to study the phenomenon of turbulence; however, the Burgers' equation does not exhibit the chaotic nature and sensitivity to initial conditions that characterizes turbulence as observed in the full and incompressible Navier-Stokes equations.

The Burgers' equation is also the simplest nonlinear extension of the linear convection-diffusion equation, and exact solutions can sometimes be found using the method of characteristics. In the scope of this dissertation, Burgers shall be used as to test the extension of our numerical method to nonlinear problems.

@@

\begin{figure}[!h]
\centering
\includegraphics[scale = .3]{figs/convCirc.png}
\caption{Setup for the vortex problem.}
\label{fig:convCirc}
\end{figure}
Formally speaking, the solution to the vortex problem is taken to be the solution to the convection-diffusion equation (with appropriate outflow boundary conditions) as the viscosity tends towards zero, in which case, the solution in the interior would be uniformly zero (this technique is referred to in mathematical literature as the ``vanishing viscosity" method, and is used to define unique solutions in the inviscid limit).  This motivates the need for \emph{artificial viscosity} methods with which to regularize inviscid problems.  The topic is expansive, and we direct the reader towards \cite{Barter} for a more detailed discussion of past and present artificial viscosity methods.

### 2013-07-22 docs/Jesse/NavierStokes paper/proposalIntro.tex
updating confusion, dissertation, and NS

\begin{figure}[!h]
\centering
\subfigure[SUPG and standard FEM solutions]{\includegraphics[scale=.43]{figs/SUPG.png}}
\subfigure[SUPG test function]{\includegraphics[scale=.22]{figs/SUPGtest.png}}
\caption{SUPG and standard Bubnov-Galerkin solutions to the 1D convection-dominated diffusion equation, and a modified SUPG test function (in black) corresponding to a linear basis ``hat" function (in red).  The upwind portion of the element is emphasized, while the downwind portion is decreased.  The magnitude of the discontinuity between the upwind and downwind portion is controlled by the intrinsic timescale parameter $\tau$. }
\label{fig:SUPG}
\end{figure}

The idea of emphasizing the upwind portion of a test function is an old idea, introduced in 1977 by Zienkiewicz et al.\ in \cite{zienkUpwind}.  However, the precise amount of upwinding,\footnote{Insufficient upwinding results in a method which still exhibits oscillations and instabilities, while excessive upwinding leads to an overly diffusive method.} as well as the connection to residual-based stabilization methods, were novel to SUPG.

>>>>

\begin{figure}[!h]
\centering
\subfigure[SUPG and standard FEM solutions]{\includegraphics[scale=.475]{figs/SUPG.png}}
\subfigure[SUPG test function]{\includegraphics[scale=.25]{figs/SUPGtest.png}}
\caption{SUPG and standard Bubnov-Galerkin solutions to the 1D convection-dominated diffusion equation, and a modified SUPG test function (in black) corresponding to a linear basis ``hat" function (in red).  The upwind portion of the element is emphasized, while the downwind portion is decreased.  The magnitude of the discontinuity between the upwind and downwind portion is controlled by the intrinsic timescale parameter $\tau$. }
\label{fig:SUPG}
\end{figure}

The idea of emphasizing the upwind portion of a test function is an older idea, introduced in 1977 by Zienkiewicz et al.\ in \cite{zienkUpwind}.  However, the precise amount of upwinding,\footnote{Insufficient upwinding results in a method which still exhibits oscillations and instabilities, while excessive upwinding leads to an overly diffusive method.} as well as the connection to residual-based stabilization methods, were novel to SUPG.
